---
title: "Germination Experiment"
author: "Mario"
date: "07/12/2021"
output: html_document
---
Purpose:
The purpose of this experiment is to determine if cooler temperatures, as seen under shrub canopies, could be used as an indicator for the germination success of both native and exotic plant species in arid ecosystems. 

Hypothesis:
Temperature can directly predict the germination success of native and exotic species.

Predictions:
1) Higher temperatures will result in a lower germination rate of native plant species.
2) Invasive plant species will have a higher germination rate under all temperature treatments.
3) Higher temperatures will have a negative effect on native and invasive species biomass.

Data:
All data can be access on KNB.
https://knb.ecoinformatics.org/view/doi:10.5063/F1GQ6W6R


```{r}
###Packages for future work
library(ggmap)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(MASS)
library(ggpubr) 
library(emmeans)
```
```{r}
##Prep Temp Data
Layia <- read.csv("All Temp Data.csv")
Brome <- read.csv("Brome Temp Sheet.csv")
Salvia <- read.csv("Salvia Temp Sheet.csv")
Phacelia <- read.csv("Phacelia Temp Sheet.csv")

Temp <- merge(Layia, Brome, all = TRUE)
Temp <- merge(Temp, Salvia, all = TRUE)
Temp <- merge(Temp, Phacelia, all = TRUE)

write.csv(Temp, "Temp.csv")

###Clean up Temperature and determine the mean and max for each pendant.
Temp2 <- read.csv("Temp2.csv")

```

```{r}
#Prep Germination Data and Combine With Temp Data
Germ <- read.csv("Final Germination.csv")
final <- merge(Germ, Temp2, all = TRUE)
```


###Data Viz
```{r}
###Temperature vs Number of Germinated Individuals by species
TempFactor <- ggplot(final, aes(Temp, number.germinated),show.legend=FALSE) +
  geom_boxplot() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempFactor
```
```{r}
###General Figure showing Temp vs total germination (Not useable)
Temp <- ggplot(final, aes(mean_temp, number.germinated),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
Temp
```

```{r}
###Figure showing Temp vs germination by species (Not facetted)
Tempspecies <- ggplot(final, aes(mean_temp, number.germinated, color = species),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
Tempspecies
```
```{r}
###This is Temperature vs germination, facet by species. I am not the biggest fan of this figure though
TempspeciesFacet <- ggplot(final, aes(mean_temp, number.germinated, color = species),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~Temp)+
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempspeciesFacet <- TempspeciesFacet + theme(legend.title = element_text(size = 3), 
               legend.text = element_text(size = 3))
TempspeciesFacet
```


```{r}
###Plot for Temperature and germination by species
TempspeciesFacet <- ggplot(final, aes(mean_temp, number.germinated),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempspeciesFacet
```

```{r}
###Statistics For Germination of Each species at varying temperatures
m1 <- glm(number.germinated ~ Temp*species+mean_temp, family = "quasipoisson", data = final)
anova(m1, test = "Chisq")
e1 <- emmeans(m1, pairwise~Temp)
e1
e2 <- emmeans(m1, pairwise~Temp:species) ###Low Temp brome germinates the same amount as Layia at all temps
e2
```
```{r}
###Mass by Temperature 
Mass <- ggplot(final, aes(mean_temp, Mass),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  theme() +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Mean Temperature", y = "Mass")
Mass
```
```{r}
MassSpecies <- ggplot(final, aes(mean_temp, Mass, color = species),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Mean Temperature", y = "Mass (g)")
MassSpecies
```
```{r}
MassSpeciesFacet <- ggplot(final, aes(mean_temp, Mass),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
MassSpeciesFacet
```

```{r}
m2 <- glm(Mass ~ number.germinated*Temp*species+mean_temp, family = "quasipoisson", data = final)
anova(m2, test = "Chisq")
e3 <- emmeans(m2, pairwise~Temp:species)
e3

```

