---
title: "Germination Experiment"
author: "Mario"
date: "07/12/2021"
output: html_document
---
Purpose:
The purpose of this experiment is to determine if cooler temperatures, as seen under shrub canopies, could be used as an indicator for the germination success of both native and exotic plant species in arid ecosystems. 

Hypothesis:
We hypothesize that variations in fine scale temperature can act as a direct proxy for successful desert plant establishment and germination.

Predictions:
1) Fine scale temperature can be experimentally manipulated via heat lamps in an enclosed setting.
2) Plants will respond to varying fine scale temperatures.
3) The response to temperature will be species specific.
4) Responses to temperature will vary between native and exotic plant species.

Data:
All data can be access on KNB.
https://knb.ecoinformatics.org/view/doi:10.5063/F1GQ6W6R

Rough Ideas:
1)To connect this to shrub density I want to show that the cooler temperatures, as experienced under shrubs, can provide a benefit for species germination. The law table is meant to simulate these lower temperatures while the medium and high tables are meant to act as open areas at a moderate temperature and at extreme temperature.
2)There should be a visible difference between overall germination and mass between native and invasive species. If we can determine if invasives can germinate better under shrubs than in open areas then we can connect this to competition between natives and invasive.
3)Species should vary in their response to fine scale temperature since the range that they germinate vary from species to species. We should see that some natives germinate better at specific temperature while others do not.
4) We can connect this possibly to climate change. If we find that higher temperatures means a lower germination success then that can be the take-home message, that increasing temperatures have negative effects on desert plant species.



```{r}
###Packages for future work
library(ggmap)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(MASS)
library(ggpubr) 
library(emmeans)
```
```{r}
##Prep Temp Data
Layia <- read.csv("All Temp Data.csv")
Brome <- read.csv("Brome Temp Sheet.csv")
Salvia <- read.csv("Salvia Temp Sheet.csv")
Phacelia <- read.csv("Phacelia Temp Sheet.csv")

###Merge all species temp data
Temp <- merge(Layia, Brome, all = TRUE)
Temp <- merge(Temp, Salvia, all = TRUE)
Temp <- merge(Temp, Phacelia, all = TRUE)

###Set up SD and SE
se_temp <- sd(Temp$temperature)/sqrt(length(Temp$temperature))
Temp$se <- se_temp

###Get mean and max temp
Temp <- Temp %>%
  group_by(as.character(temp), species, pendant_ID, se) %>%
  summarise(mean_temp = mean(temperature), max_temp = max(temperature))
names(Temp)[1] <- "temp"

###Output file
write.csv(Temp, "Temp.csv")

```

```{r}
###Clean up Temperature and determine the mean and max for each pendant.
Temp2 <- read.csv("Temp_2.csv")
```


```{r}
#Prep Germination Data and Combine With Temp Data
Germ <- read.csv("Final Germination.csv")
final <- merge(Germ, Temp2, all = TRUE)

write.csv(final, "final.csv")
```


###Data Viz
```{r}
###Temperature vs Number of Germinated Individuals in 6 weeks by species
TempFactor <- ggplot(final, aes(temp, germination),show.legend=FALSE) +
  geom_boxplot() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() + labs(tag = "A")+
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempFactor
```
```{r}
###Plotting the lm
plot(lm(germination ~ mean_temp, data = final), which = 1)
```
```{r}
germination_poly <- lm(germination ~ mean_temp + I(mean_temp^2), data = final)

library(car)
library(broom)
Anova(germination_poly)
```
```{r}
###General Figure showing Temp vs total germination in 6 weeks (Not useable)
Temp <- ggplot(final, aes(mean_temp, germination),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
Temp
```

```{r}
###Figure showing Temp vs germination by species (Not facetted)
Tempspecies <- ggplot(final, aes(mean_temp, germination, color = species),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
Tempspecies
```
```{r}
###This is Temperature vs germination, facet by species. I am not the biggest fan of this figure though
TempspeciesFacet <- ggplot(final, aes(mean_temp, germination, color = species),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~temp)+
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempspeciesFacet <- TempspeciesFacet + theme(legend.title = element_text(size = 3), 
               legend.text = element_text(size = 3))
TempspeciesFacet
```


```{r}
###Plot for Temperature and germination in 6 weeks by species
TempspeciesFacet <- ggplot(final, aes(mean_temp, germination),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempspeciesFacet
```
```{r}
###Figure shows temperature v Germination in 6 weeks, facet by species and filled via temperature
TempspeciesFacetColor <- ggplot(final, aes(mean_temp, germination, color = temp),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Number of Germinated Seeds")
TempspeciesFacetColor
```


```{r}
###Statistics For Germination of Each species at varying temperatures
m1 <- glm(germination ~ temp*species+mean_temp, family = "quasipoisson", data = final)
anova(m1, test = "Chisq")
###Was temp manipulated significantly in the experiment?
e1 <- emmeans(m1, pairwise~temp)
e1

###Germination by temp and species
e2 <- emmeans(m1, pairwise~temp|species) ###Low Temp brome germinates the same amount as Layia at all temps
e2
```
```{r}
###Difference in germination by species and temp
e3 <- emmeans(m1, pairwise~species|temp) ###Low Temp brome germinates the same amount as Layia at all temps
e3
```


```{r}
###Mass by Temperature 
Mass <- ggplot(final, aes(mean_temp, mass),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  theme() +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Mean Temperature", y = "Mass")
Mass
```
```{r}
MassSpecies <- ggplot(final, aes(mean_temp, mass, color = species),show.legend=FALSE) +
  geom_point() +
  scale_color_brewer(palette = "Set1") + theme_classic() + 
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Mean Temperature", y = "Mass (g)")
MassSpecies
```
```{r}
MassSpeciesFacet <- ggplot(final, aes(mean_temp, mass),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() +
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Mass (g)")
MassSpeciesFacet
```
```{r}
###Temperature v biomass, facet by species and fill by Temp
MassSpeciesFacetColor <- ggplot(final, aes(mean_temp, mass, color = temp),show.legend=FALSE) +
  geom_point() +
  facet_wrap(~species)+
  scale_color_brewer(palette = "Set1") + theme_classic() +
  theme(axis.title.x = element_blank()) +
  geom_smooth(method = lm, se = TRUE) + 
  labs(x = "Temperature", y = "Biomass (g)")
MassSpeciesFacetColor
```


```{r}
###Biomass by temp and species
m2 <- glm(mass ~ germination*temp*species+mean_temp, family = "quasipoisson", data = final)
anova(m2, test = "Chisq")
e4 <- emmeans(m2, pairwise~temp|species)
e4

```
```{r}
###Difference in biomass by species and temp
e5 <- emmeans(m2, pairwise~species|temp)
e5

```
